---
title: "Homework5 Models in R"
format: pdf
editor: visual
---

```{r, include=FALSE}
rm(list=ls())
```

## Task One: Conceptual Questions

### Question One: What is the purpose of using cross-validation when fitting a random forest model?

### Question Two: Describe the bagged tree algorithm.

### Question Three: What is meant by a general linear model?

### Question Four: When fitting a multiple linear regression model, what does adding an interaction term
do? That is, what does it allow the model to do differently as compared to when it is
not included in the model?

### Question Five: Why do we split our data into a training and test set?


## Task Two: Data Prep

### Packages and data

```{r, warning=FALSE, message=FALSE}
library("tidyverse")
library("tidymodels")
library("caret")
library("yardstick")

heartdata <- read.csv("heart.csv")
```

### Question One

```{r}
summary(heartdata)
```

#### a. What type of variable (in R) is Heart Disease? Categorical or Quantitative?

Heart Disease is a numerical (integer) variable. It is quantitative rather than categorical. We know this because a numerical summary is presented for the Heart Disease variable. 

#### b. Does this make sense? Why or why not.

This does not make sense because a person has either been diagnosed with heart disease or they have not. 

### Question Two

```{r}
new_heart <- heartdata |>
  mutate(new_hd = as.factor(HeartDisease)) |>
  select(-c(ST_Slope, HeartDisease))

summary(new_heart)
```

